Theoretical prediction of disrupted Min oscillation in flattened \emph{Escherichia coli}

### 2015-06-27 paper/plos.tex
changes

* E-mail: Corresponding schuljef@physics.oregonstate.edu
\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
  The dynamics of the Min-protein system help \emph{Escherichia coli}
  regulate the process of cell division by identifying the center of
  the cell.  While this system exhibits robust bipolar oscillations in
  wild-type cell shapes, recent experiments have shown that when the
  cells are mechanically deformed into wide, flattened out, irregular
  shapes, the spatial regularity of these oscillations breaks down. We
  employ widely used stochastic and deterministic models of the Min
  system to simulate cells with flattened shapes.  The deterministic
  model predicts strong bipolar oscillations, in contradiction with
  the experimentally observed behavior, while the stochastic model,
  which is based on the same reaction-diffusion equations, does
  predict spatially irregular oscillations as observed in experiment.
  We further report simulations of flattened but more symmetric
  shapes, and find that it is the flattening and lateral expansion
  rather than the asymmetry of the cell shapes that causes the
  irregular oscillation behavior.

@@

However Fig.~\ref{randst-plot-ave} leaves unclear the importance of
the flattening and accompanying lateral expansion of the cell versus
cell shape irregularity and asymmetry in the temporal regularity of
the oscillations.  Although in the movies the stadiums appear to
display, on average over long time scales, a somewhat more regular
oscillation than appears in the irregular shapes \emph{A} and
\emph{B}, the limited time range of the arrow plots (350 seconds) is
inadequate to make this comparison.  We therefore return to the
correlation function between the number of proteins in a segment at
the top of the cell and the number of proteins in a segment at the
bottom of the cell.  Fig.~\ref{fig:corr-B} shows this plot for shape
\emph{B} and stadium \emph{B}.  Here we plot correlations of
correlation times up to 500 seconds, a biologically relevant
timescale.  E. Coli cell doubling times vary widely depending on
environmental conditions, but will often be between 20 minutes to 100
minutes~\cite{pierucci1972chromosome}, while fluorescent microscopy
experiments have shown that the FtsZ ring builds itself with a half
life of roughly 30 seconds~\cite{stricker2002rapid}.

@@

\section*{Conclusion}
We find agreement between our stochastic simulations and experiments
showing disrupted bipolar MinD oscillation in asymmetric flattened
E. coli cells~\cite{mannik2012robustness}.  As observed
experimentally, our simulations predict that MinD maxima will form in
a spatially irregular sequence of locations.  This result builds upon
existing results showing that this model and its variants are
effective in a variety of wild-type and mutated cell
shapes~\cite{fange2006noise, varma2008min, kruse2007experimentalist},
and reinforces that stochastic variations of Huang's 2003
model~\cite{fange2006noise, kerr2006division} has considerable
predictive power beyond the standard wild-type cell.

>>>>

* E-mail: Corresponding schuljef@physics.oregonstate.edu
\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
  The dynamics of the Min-protein system help \emph{Escherichia coli}
  regulate the process of cell division by identifying the center of
  the cell.  While this system exhibits robust bipolar oscillations in
  wild-type cell shapes, recent experiments have shown that when the
  cells are mechanically deformed into wide, flattened out, irregular
  shapes, the spatial regularity of these oscillations breaks down. We
  employ widely used stochastic and deterministic models of the Min
  system to simulate cells with flattened shapes.  The deterministic
  model predicts strong bipolar oscillations, in contradiction with
  the experimentally observed behavior, while the stochastic model,
  which is based on the same reaction-diffusion equations, does
  predict spatially irregular oscillations as observed in experiment\fixme{, although they are generally more bipolar}.
  We further report simulations of flattened but more symmetric
  shapes, and find that it \fixme{may be?} is the flattening and lateral expansion
  rather than the asymmetry of the cell shapes that causes the
  irregular oscillation behavior.

@@

However Fig.~\ref{randst-plot-ave} leaves unclear the importance of
the flattening and accompanying lateral expansion of the cell versus
cell shape irregularity and asymmetry in the temporal regularity of
the oscillations.  Although in the movies the stadiums appear to
display, on average over long time scales, a somewhat more regular
oscillation than appears in the irregular shapes \emph{A} and
\emph{B}, the limited time range of the arrow plots (350 seconds) is
inadequate to make this comparison.  We therefore return to the
correlation function between the number of proteins in a segment at
the top of the cell and the number of proteins in a segment at the
bottom of the cell.  \fixme{this is where Ill need to add the volumes} Fig.~\ref{fig:corr-B} shows this plot for shape
\emph{B} and stadium \emph{B}.  Here we plot correlations of
correlation times up to 500 seconds, a biologically relevant
timescale.  E. Coli cell doubling times vary widely depending on
environmental conditions, but will often be between 20 minutes to 100
minutes~\cite{pierucci1972chromosome}, while fluorescent microscopy
experiments have shown that the FtsZ ring builds itself with a half
life of roughly 30 seconds~\cite{stricker2002rapid}.

@@

\section*{Conclusion}
\fixme{We find a certain amount of qualitive agreement between our stochastic simulations and experiments showing disrupted MinD oscillations in asymmetric flattened E. coli cells.}We find agreement between our stochastic simulations and experiments
showing disrupted bipolar MinD oscillation in asymmetric flattened
E. coli cells~\cite{mannik2012robustness}.  As observed
experimentally, our simulations predict that MinD maxima will form in
a spatially irregular sequence of locations, \fixme{albeit with a with more course scale bipolar regularity.}  This result builds upon
existing results showing that this model and its variants are
effective in a variety of wild-type and mutated cell
shapes~\cite{fange2006noise, varma2008min, kruse2007experimentalist},
and reinforces that stochastic variations of Huang's 2003
model~\cite{fange2006noise, kerr2006division} has considerable
predictive power beyond the standard wild-type cell.
